---
title: "Appendix "
output: html_document
---

# Section A.1 Database Control Tables

**Table A-1 MAZ (ARC TAZ) Level Control Totals**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}
library(tidyverse)
library(knitr)
library(DT)
library(kableExtra)
library(readxl)

A1<- read_excel('data/UserGuideTables.xlsx','A1','B3:D22')
A1[is.na(A1)] <- ''

kable(A1) %>%
  kable_styling("striped")

```

**Table A-2 TAZ (ARC PECAS Zone) Level Control Totals**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A2<- read_excel('data/UserGuideTables.xlsx','A2','B3:D10')
A2[is.na(A2)] <- ''

kable(A2) %>%
  kable_styling("striped")

```


**Table A-3 Meta-Control (ARC County) Totals**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A3<- read_excel('data/UserGuideTables.xlsx','A3','B3:D14')
A3[is.na(A3)] <- ''

kable(A3) %>%
  kable_styling("striped")

```

# Section A.2 Database PUMS Tables

**Table A-4 PUMS Processed Household Table**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A4<- read_excel('data/UserGuideTables.xlsx','A4','B3:D7')
A4[is.na(A4)] <- ''

kable(A4) %>%
  kable_styling("striped")

```

**Table A-5 PUMS Processed Person Table**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A5<- read_excel('data/UserGuideTables.xlsx','A5','B3:D9')
A5[is.na(A5)] <- ''

kable(A5) %>%
  kable_styling("striped")

```


# Section A.3 Database Output Tables

**Table A-6 Synthetic Population Household Table in Unexpanded Form**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A6<- read_excel('data/UserGuideTables.xlsx','A6','B3:C14')
A6[is.na(A6)] <- ''

kable(A6) %>%
  kable_styling("striped")

```


**Table A-7 Synthetic Population Person Table in Unexpanded Form**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A7<- read_excel('data/UserGuideTables.xlsx','A7','B3:C15')
A7[is.na(A7)] <- ''

kable(A7) %>%
  kable_styling("striped")

```


**Table A-8 Synthetic Population Household Table in Expanded Form**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A8<- read_excel('data/UserGuideTables.xlsx','A8','B3:C15')
A8[is.na(A8)] <- ''

kable(A8) %>%
  kable_styling("striped")

```

**Table A-9 Synthetic Population Person Table in Expanded Form**    

```{r,echo=F,message=F,warning=F,output=F,out.height=700}

A9<- read_excel('data/UserGuideTables.xlsx','A9','B3:C17')
A9[is.na(A9)] <- ''

kable(A9) %>%
  kable_styling("striped")

```


Notes: The definitions for other PUMS variables can be found in the 2007-2011 ACS PUMS Data Dictionary:   
$^{*}$ http://www2.census.gov/programs-surveys/acs/tech_docs/pums/data_dict/PUMS_Data_Dictionary_2007-2011.pdf   